---
title: "R Mini Tutorial: Basics objects"
author: "luiscartor"
date: "May 29, 2019"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Getting started

In Unix or Mac, execute R on the command line, or click on the R icon on the desktop for Windows. 

To finish an R session, type `q()` at the command prompt. You will be asked to save the workspace when finishing the session. It is a good practice to NOT to save the session, but if you say yes, the objects on your workspace will be saved as a .RData file. If saved, these workspace will be loaded when you open a new session. 

## 1.1 Basic calculations 

Use the console to execute R commands, such as basic calculations. After hiting enter, the console will show the result.

```{r}
  3+2
```

But the result of the calculation can be also stored to a variable:

```{r}
  a <- 3+2
```

Variable names in R must begin with a letter, followed by aphanumeric characters. Periods or underscores are often used to break up long variables (e.g. `long_variable`). Use variable names that help you identify the variable, but without being very long (e.g. herons_popmean). Avoid `c`, `q`, `t`, `C`, `D`, `F`, `I`, `T`, and things like `mean`, `var`, `sd`, as they are built-in R functions and they will cause errors. 

Now '5' is stored in the variable a. To print the value of a variable, just type the variable name:

```{r}
  a
```

Some arithmetic operators:

|Operator | Description |
|--- | --- |
|+| 	addition|
|-| 	subtraction|
|*| 	multiplication|
|/| 	division|
|^ or ** 	|exponentiation|

Some comparison and logical operators:

|Operator |	Description|
|--- | --- |
|> |	greater than|
|>= |	greater than or equal to|
|== |	exactly equal to|
|!= |	not equal to| 
|! |	logical NOT|
|& |	logical AND|


Now we can start creating variables and making arithmetic and logical operations. 

```{r}
x <- 5
y <- 2
z1 <- x*y
z2 <- x/y
z3 <- x^y
z1; z2; z3
```


Note that we can use a semicolon to put two or more commands in one line. In the last line of code, we use the semicolon to print the values of the three z variables.

Apart from
|sqrt() | squared root |


We can combine several operations:

```{r}
A <- 5; B <- 3
C <- (A+B)/(A+5*B) 
C
```

And use the logical operators:

```{r}
A == B
```
```{r}
A > B
```
```{r}
C != B
```


We can use a wide rage of built-in mathematical functions in R

|R command |	function|
|--- | --- |
|abs() |	absolute value, $|
|cos(), sin(), tan() |	cosine, sine, tangent|
|acos(), asin(), atan() |	arc-cosine, arc-sine, arc-tangent|
|exp(x) |	exponential function|
|log() |	natural (base-e) logarithm|
|log10() |	base-10 logarithm|
|sqrt() |	square root, √x|

For example, try this calculations:

```{r}
A <- 5; B <- 3
C <- (A+2*sqrt(B))/(A+5*sqrt(B)) 
D <- log(A+B)/cos(C)
C;D
```


As usual, parentheses alter the order of the operations. Try:

```{r}
C <- A+2*sqrt(B)/A+5*sqrt(B)
C
```

and

```{r}
C <- A + 2*(sqrt(B)/A + 5) *sqrt(B)
C
```

<br>

> ### Excersise: Solve the following equations
>
> 1. $$1+0.3+0.7^2$$
> 2. For a = 0.5 and b = 1: $$\frac{sin(a)cos(b)}{cos(a)}$$ 
> 3. $$e^{0.5^2/4}$$
> 4. For a = 10 and b = 2: $$ \log(a)-3 * \log10(b) $$

<br>
$$\frac{sin(a)cos(b)}{cos(a)}$$ 

## 1.2 Basics Data Types

## 1.2.1 Vectors and matrices



```{r}
  vec <- seq(1:5)
  vec[1:3]
```

















```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
